Spherical Harmonics and Vector Spherical Harmonics as Orthogonal Polynomials

### 2018-07-06 spherical-harmonics-paper/presentation-spherical-harmonics.tex
Second draft

Let \((x,y,z)\) be a point on the unit sphere such that \(x^2+y^2+z^2 = 1\). Let \(\theta, \varphi\) denote the angles such that
\begin{align}
x &= \sintheta \cosphi \\
y &= \sintheta \sinphi \\
z &= \costheta.
\end{align}
Further, define
\begin{align}
\clm := \Bigg(\frac{(2l+1)\Gamma(l-m+1)}{4\pi\Gamma(l+m+1)}\Bigg)^\frac{1}{2},
\end{align}
and
\begin{align}
\alphalm := \begin{cases} 
		\clm \chatmod \quad \quad \quad \text{if } m \ge 0 \\
		\clm \chatmod \ctildemod \quad \text{if } m < 0
	   \end{cases}.
\end{align}

@@

The Clenshaw algorithm is then as follows:
\begin{align*}
\quad &\text{1) } \text{Set } \bm{\gamma}_{N+2} = \bold{0}, \: \bm{\gamma}_{N+2} = \bold{0}. \\
\quad &\text{2) } \text{For } n = N:-1:1 \\
\quad & \quad \quad \quad \text{set } \bm{\gamma}_{n}^T = \bold{f}_n^T - \bm{\gamma}_{n+1}^T D^T_n (B_n - G_n) -  \bm{\gamma}_{n+2}^T D^T_{n+1}C_{n+1} \\
\quad &\text{3) } \text{Output: } f(x,y,z) \approx \bigP_0(x,y,z) f_0 + \bm{\gamma}_{1}^T \bigP_1(x,y,z) - \bigP_0(x,y,z) \bm{\gamma}_{2}^T D^T_1 C_1.
\end{align*}

@@

\frame
{
    \frametitle{Vector Spherical Harmonics}

\item In general, the tensor spherical harmonic is given by, for \(2l, \, j, \, 2s \in \No\),
\begin{align}
\curlyy^{j,s}_{l,m}(x,y,z) &= \sum_{m_s=-s}^s \langle j \quad m-m_s \; ; \; s \quad m_s \; | \; l \quad m \rangle \: \: Y^{m-m_s}_j (x,y,z) \: \: \bm{\chi}_{s, m_s},
\end{align}

\item \(\bm{\chi}_{s, m_s}\) are the simultaneous eigenstates of the spin operators \(\bold{S}^2\) and \(S_z\).

\item \(\langle j \quad m-m_s \; ; \; s \quad m_s \; | \; l \quad m \rangle\) is a Clebsch-Gordan coefficient. 
\begin{itemize}
	\item C-G coefficients vanish unless \(|j - s| \le l \le j + s\). 
	\item There are simple calculable expressions for the C-G coefficients when the spin \(s=1\).
\end{itemize}

\item Both \(\gradYlm\), \(\gradpYlm\) can be expressed in terms of spin-1 tensor spherical harmonics (vector SHs), which in turn can each be expressed as a vector-weighted sum of spherical harmonics.

\frame
{
    \frametitle{Vector Spherical Harmonics}

To simplify the notation, for spin $s=1$, we define
\begin{align}
\cgcoeff^{L, m_s}_{l, m} := \, \langle L \quad m-m_s \; ; \; 1 \quad m_s \; | \; l \quad m \rangle.
\end{align}

Thus we have that the three spin-1 tensor spherical harmonics are given by
\begin{align}
\curlyy^{l \pm1,s}_{l,m}(x,y,z) &= \sum_{m_s=-1}^1 \cgcoeff^{l\pm1, m_s}_{l, m} \, Y^{m-m_s}_{l \pm 1} (x,y,z) \, \chivec, \\
\curlyy^{l,s}_{l,m}(x,y,z) &= \sum_{m_s=-1}^1 \cgcoeff^{l, m_s}_{l, m} \, Y^{m-m_s}_l (x,y,z) \, \chivec.
\end{align}
where
\begin{align}
\bm{\chi}_{1, \pm 1} = \frac{1}{\sqrt{2}} \begin{bmatrix} \mp 1 \\ -i \\ 0 \end{bmatrix}, \quad \bm{\chi}_{1,0} = \begin{bmatrix} 0 \\ 0 \\ 1 \end{bmatrix}.
\end{align}

@@

@@

Operators act on the coefficients vectors.

Easily derive operators for Grad, Div, curl with the unit vector, etc. as well as the multiplications by $x, y, z$.

- BandedBlockBanded structures

Can calculate dot products simply.

>>>>

\begin{block}{Motivation}
To look at creating a library for working with spherical harmonics (SHs) as orthogonal polynomials (OPs) in three variables on the unit sphere.
\end{block}

Let \((x,y,z)\) be a point on the unit sphere, \(\theta, \varphi\) denote the angles such that
\begin{align}
x &= \sintheta \cosphi \\
y &= \sintheta \sinphi \\
z &= \costheta.
\end{align}
Further, define
\begin{align}
\clm := \Bigg(\frac{(2l+1)\Gamma(l-m+1)}{4\pi\Gamma(l+m+1)}\Bigg)^\frac{1}{2},
\end{align}
and
\begin{align}
\alphalm := \begin{cases} 
		\clm \chatmod \quad \quad \quad \text{if } m \ge 0 \\
		\clm \chatmod \ctildemod \quad \text{if } m < 0
	   \end{cases}.
\end{align}

@@

The Clenshaw algorithm is then as follows:
\begin{align*}
\quad &\text{1) } \text{Set } \bm{\gamma}_{N+2} = \bold{0}, \: \bm{\gamma}_{N+2} = \bold{0}. \\
\quad &\text{2) } \text{For } n = N:-1:1 \\
\quad & \quad \quad \quad \text{set } \bm{\gamma}_{n}^T = \bold{f}_n^T - \bm{\gamma}_{n+1}^T D^T_n (B_n - G_n(x,y,z)) -  \bm{\gamma}_{n+2}^T D^T_{n+1}C_{n+1} \\
\quad &\text{3) } \text{Output: } f(x,y,z) \approx \bigP_0(x,y,z) f_0 + \bm{\gamma}_{1}^T \bigP_1(x,y,z) - \bigP_0(x,y,z) \bm{\gamma}_{2}^T D^T_1 C_1.
\end{align*}

@@

@@

\begin{itemize}
\item $\curlyy^{l-1,1}_{l,m}, \curlyy^{l,1}_{l,m}, \curlyy^{l+1,1}_{l,m}$ are spin-1 tensor spherical harmonics (vector SHs).
\item VSHs can be expressed as a vector-weighted sum of spherical harmonics.
\end{itemize}

@@

\item Operators act on the coefficients vectors.

\item Easily derive operators for Grad, Div, curl with the unit vector, etc. as well as the multiplications by $x, y, z$.
\begin{itemize}
	\item BandedBlockBanded structures
\end{itemize}

\item Can calculate dot products simply.
